---
title: "Instacart Exploratory Analysis"
output:
  html_document: default
  html_notebook: default
---


```{r}
source("readDataToMemory.R")
readInstacart()

library(DBI)
library(ggplot2)
library(ggthemes)

src_tbls(sc)
```

### Order products

`order_products__*.csv`

These files specify which products were purchased in each order.

```{r}
dbGetQuery(sc, "SELECT * FROM order_products__prior_tbl LIMIT 10")
```
 
## orders

`orders.csv`

This file tells to which set (prior, train, test) an order belongs. 


```{r}
orders_prev <- dbGetQuery(sc, "SELECT * FROM orders_tbl LIMIT 100")
orders_prev
```
 
```{r}
dbGetQuery(
  sc,
  "
SELECT COUNT(DISTINCT product_id) AS products
FROM order_products__prior_tbl
  "
  )
```

```{r}
products_popularity_txt <- "
SELECT product_id
,   COUNT(1) AS n_orders
FROM order_products__prior_tbl
GROUP BY product_id
ORDER BY n_orders DESC
LIMIT 50
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Products

`products.csv`

```{r}
dbGetQuery(sc, "SELECT * FROM products_tbl LIMIT 20")
```

## Joins

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

```{r}

products_popularity %>%
  ggplot(
    aes(reorder(product_name, n_orders, function(x) x), 
        n_orders)) +
  geom_bar(stat="identity", fill='darkorange') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("product") +
  ylab("Number of orders") +
  theme_minimal()
```

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Questions

* Which day of week people order most fequently? Day 0 (Sunday?)
```{r}
day_popularity_txt <- "
SELECT order_dow
,   COUNT(1) AS n_orders
FROM orders_tbl
GROUP BY order_dow
ORDER BY n_orders DESC
LIMIT 50
"

day_popularity <-
  dbGetQuery(sc, day_popularity_txt)

day_popularity
```
* The most popular hour of doing order? 10AM
```{r}
hour_popularity_txt <- "
SELECT order_hour_of_day
,   COUNT(1) AS n_orders
FROM orders_tbl
GROUP BY order_hour_of_day
ORDER BY n_orders DESC
LIMIT 50
"

hour_popularity <-
  dbGetQuery(sc, hour_popularity_txt)

hour_popularity
```
* The most reordered products? Banana

```{r}
reorder_popularity_txt <- "
SELECT op.product_id,n_orders,product_name
FROM (
  SELECT product_id,COUNT(1) AS n_orders
  FROM order_products__prior_tbl
  WHERE reordered=1
  GROUP BY product_id
  ORDER BY n_orders DESC
  LIMIT 30)op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"


reorder_popularity <-
  dbGetQuery(sc, reorder_popularity_txt)

reorder_popularity

```

* The less reorderd product? Pickled Ginger
```{r}
reorder_popularity_txt <- "
SELECT op.product_id,n_orders,product_name
FROM (
  SELECT product_id,COUNT(1) AS n_orders
  FROM order_products__prior_tbl
  WHERE reordered=1
  GROUP BY product_id
  ORDER BY n_orders ASC
  LIMIT 30)op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"


reorder_popularity <-
  dbGetQuery(sc, reorder_popularity_txt)

reorder_popularity

```
* Plot histogram of number of products in an order.

```{r}
num_products_txt <- "
SELECT order_id,COUNT(1) AS n_products
FROM order_products__prior_tbl
GROUP BY order_id
"


num_products <-
  dbGetQuery(sc, num_products_txt)

num_products

num_products %>%
  ggplot(
    aes(reorder(order_id, n_products, function(x) x), 
        n_products)) +
  geom_bar(stat="identity", fill='darkorange') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("orders") +
  ylab("Number of products") +
  theme_minimal()

```

* What is the most popular department? 
```{r}
department_popularity_txt <- "
SELECT department_id,SUM(n_orders) AS total_orders
FROM(
  SELECT op.product_id,department_id,n_orders
  FROM (
    SELECT product_id,COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC)op
  LEFT JOIN (
      SELECT product_id,department_id
      FROM products_tbl) p
  ON op.product_id = p.product_id
ORDER BY total_orders ASC
)
"

department_popularity <-
  dbGetQuery(sc, department_popularity_txt)

department_popularity

```

## Whisker, magrittr

```{r}
library(whisker)
library(magrittr)

product_times_ordered_txt <- "
SELECT n_orders
,   COUNT(DISTINCT user_id) AS n_users
FROM (
    SELECT o.user_id
    ,   COUNT(DISTINCT op.order_id) AS n_orders
    FROM (
        SELECT order_id
        FROM order_products__prior_tbl
        WHERE product_id = {{productId}}
    ) op
    LEFT JOIN (
        SELECT  user_id
        ,   order_id
        FROM orders_tbl) o
    ON op.order_id = o.order_id
    GROUP BY o.user_id) opo
GROUP BY n_orders
ORDER BY n_orders
"

data <- list(productId = 24852)

product_times_ordered_txt %>%
  whisker.render(data) %>%
  cat
```
```{r}
product_times_ordered_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .) %>%
  ggplot(aes(n_orders, n_users)) + 
  geom_line() +
  theme_minimal()
```

## Questions

* For given department, list the most popular 10 products (with templating department).
```{r}
product_times_ordered_txt <- "
SELECT product_name,SUM(op.product_id) AS n_orders
FROM(
  SELECT op.product_id,department_id,n_orders
  FROM (
    SELECT product_id,COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC)op
  LEFT JOIN (
      SELECT product_id,department_id
      FROM products_tbl) p
  ON op.product_id = p.product_id
LIMIT 10
)
"
product_times_ordered <-
  dbGetQuery(sc, product_times_ordered_txt)

product_times_ordered

```

* For given day, rank department popularity.

## More on orders (Window Functions)



```{r}
dbGetQuery(sc, 
"
SELECT days_since_prior_order
,   COUNT(1) AS N
FROM orders_tbl
GROUP BY days_since_prior_order
") 
```


```{r}
dbGetQuery(sc, 
"SELECT * 
FROM orders_tbl
WHERE user_id <= 10
")
```
```{r}
dbGetQuery(sc, 
"SELECT user_id,order_id,FIRST_VALUE(order_id)  OVER(partition by user_id ORDER BY order_number) AS first_order
FROM orders_tbl
WHERE user_id <= 10 
")
```



```{r}
dbGetQuery(sc, 
"SELECT product_id
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn
GROUP BY product_id
ORDER BY avg_products_number DESC
LIMIT 100
")
```




* Can you add the name? YES

```{r}
dbGetQuery(sc, 
"SELECT product_name,AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id,product_id,COUNT(product_id) OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) op
LEFT JOIN (
    SELECT product_id,product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
GROUP BY product_name
ORDER BY avg_products_number DESC
LIMIT 10

")
```

* What product is ordered when the order is small? Laxative

```{r}
dbGetQuery(sc, 
"SELECT product_name,AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id,product_id,COUNT(product_id) OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) op
LEFT JOIN (
    SELECT product_id,product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
GROUP BY product_name
ORDER BY avg_products_number ASC
LIMIT 10

")
```

* For 10 first users and all their orders find days since first order

```{r}

```



